Students' Alcohol Consumption impacts on academic Results on the example of a secondary Scool Maths Class in Portugal

### 2016-10-13 MSproject.Rmd
Remi codes

>>>>

```{r pressure, echo=FALSE}
plot(pressure)
```
```{r}
sm <- student.mat
for (i in 16:24) {                                  
  sm[,i] <- as.character(sm[,i])
  for (j in 1:nrow(sm)) {
    if (sm[j,i] == "yes") sm[j,i] <- 1
    if (sm[j,i] == "no") sm[j,i] <- 0
  }
  sm[,i] <- as.numeric(sm[,i])
}

for (i in 6) {                                  
  sm[,i] <- as.character(sm[,i])
  for (j in 1:nrow(sm)) {
    if (sm[j,i] == "A") sm[j,i] <- 1
    if (sm[j,i] == "T") sm[j,i] <- 0
  }
  sm[,i] <- as.numeric(sm[,i])
}

for (i in 5) {                                  
  sm[,i] <- as.character(sm[,i])
  for (j in 1:nrow(sm)) {
    if (sm[j,i] == "GT3") sm[j,i] <- 1
    if (sm[j,i] == "LE3") sm[j,i] <- 0
  }
  sm[,i] <- as.numeric(sm[,i])
}

sm[,"Medu"] <- abs(sm[,"Medu"] - 6)
sm[,"Fedu"] <- abs(sm[,"Fedu"] - 6)
sm[,"famrel"] <- abs(sm[,"famrel"] - 6)
sm[,"famrel"] <- abs(sm[,"famrel"] - 6)

#---------------------------------------------------------------------

#General test 
```{r}
t.test(sm[,"Walc"], sm[,"Dalc"])
library(ggplot2)
ggplot(data = sm) + geom_bar(aes(x=age))
ggplot(data = sm) + geom_bar(aes(x=sex))
ggplot(data = sm) + geom_bar(aes(x=Dalc))
ggplot(data = sm) + geom_bar(aes(x=Walc))
i <- ggplot(sm, aes(x = sm[,"age"], y = sm[,"sex"]))
i + geom_bin2d(binwidth = c(1,1)) #LABEL #Distribution of gender and age
i <- ggplot(sm, aes(x = sm[,"Dalc"], y = sm[,"Walc"]))
i + geom_bin2d(binwidth = c(1,1)) #LABEL #Distribution of Walc and Dalc

#Test and plot on gender and age

###age
```{r}
f <- ggplot(data =sm, aes(x = sm[,"Dalc"], y = sm[,"age"]))
f + geom_smooth(method = lm) #LABEL
f <- ggplot(data =sm, aes(x = sm[,"Walc"], y = sm[,"age"]))
f + geom_density(method = lm)  #LABEL
f + geom_smooth(na.rm = TRUE)
i <- ggplot(sm, aes(x = sm[,"age"], y = sm[,"sex"]))
i + geom_bin2d(binwidth = c(1,1))
i <- ggplot(sm, aes(x = sm[,"absences"], y = sm[,"Walc"]))
i + geom_bin2d(binwidth = c(1,1))
cor.test(sm[,"age"],sm[,"Dalc"])
cor.test(sm[,"age"],sm[,"Walc"])
```

###gender
```{r}
for (i in 2) {                                  
  sm[,i] <- as.character(sm[,i])
  for (j in 1:nrow(sm)) {
    if (sm[j,i] == "M") sm[j,i] <- 1
    if (sm[j,i] == "F") sm[j,i] <- 0
  }
  sm[,i] <- as.numeric(sm[,i])
}
cor.test(sm[,"sex"],sm[,"Walc"])
cor.test(sm[,"sex"],sm[,"Dalc"])
```

##Correlations tests 
As inputs have to be numeric it requires less lines to split the code that way instead of doing block per parameters analysed (Failures, paid, Medu, ... )
We could have done for loops here to calculate several correlations but it the results wasn't user friendly and doint it that way gave some useless correlations.
```{r}
sm <- as.numeric(sm)
sm1 <- subset(sm, sex == 1) #Create subsets with boys or girls
sm1 <- subset(sm, sex == 0)
```

###Romantic
```{r}
cor.test(sm[,"romantic"], sm[,"Dalc"])
cor.test(sm[,"romantic"], sm[,"Walc"])
```

###absenses
```{r}
cor.test(sm[,"absences"], sm[,"Dalc"])
cor.test(sm[,"absences"], sm[,"Walc"])
```

###Failures
```{r}
cor.test(sm[,"failures"], sm[,"Dalc"])
cor.test(sm[,"failures"], sm[,"Walc"])
```

###Boypart
```{r}
cor.test(sm1[,"failures"], sm1[,"Dalc"])
cor.test(sm1[,"failures"], sm1[,"Walc"])
```

###Medu
```{r}
cor.test(sm[,"Medu"], sm[,"Dalc"])
cor.test(sm[,"Medu"], sm[,"Walc"])
```

###Fedu
```{r}
cor.test(sm[,"Fedu"], sm[,"Dalc"])
cor.test(sm[,"Fedu"], sm[,"Walc"])
```

###Travel Time
```{r}
cor.test(sm[,"traveltime"], sm[,"Dalc"])
cor.test(sm[,"traveltime"], sm[,"Walc"])
```

###famrel
```{r}
cor.test(sm[,"famrel"], sm[,"Dalc"])
cor.test(sm[,"famrel"], sm[,"Walc"])
```

###goout
```{r}
cor.test(sm[,"goout"], sm[,"Dalc"])
cor.test(sm[,"goout"], sm[,"Walc"])
```

###famsize
```{r}
cor.test(sm[,"famsize"], sm[,"Dalc"])
cor.test(sm[,"famsize"], sm[,"Walc"])
```

###internet
```{r}
cor.test(sm[,"internet"], sm[,"Dalc"])
cor.test(sm[,"internet"], sm[,"Walc"])
```

###famrel
```{r}
cor.test(sm[,"famrel"], sm[,"Dalc"])
cor.test(sm[,"famrel"], sm[,"Walc"])
```

###Grades
```{r}
cor.test(sm[,"G1"], sm[,"Dalc"])
cor.test(sm[,"G1"], sm[,"Walc"])
cor.test(sm[,"G2"], sm[,"Dalc"])
cor.test(sm[,"G2"], sm[,"Walc"])
cor.test(sm[,"G3"], sm[,"Dalc"])
cor.test(sm[,"G3"], sm[,"Walc"])
cor.test(sm[,"G3"], sm[,"goout"])
cor.test(sm[,"G3"], sm[,"failures"])
cor.test(sm[,"G3"], sm[,"absences"])
```

#EARLY SIGNS OF ACADEMIC FAILURES
###abscences
```{r}
cor.test(sm[,"absences"], sm[,"Dalc"])
cor.test(sm[,"absences"], sm[,"Walc"])
cor.test(sm1[,"absences"], sm1[,"Dalc"])
cor.test(sm1[,"absences"], sm1[,"Walc"])
cor.test(sm2[,"absences"], sm2[,"Dalc"])
cor.test(sm2[,"absences"], sm2[,"Walc"])
```

###Failures
```{r}
cor.test(sm[,"failures"], sm[,"Dalc"])
cor.test(sm[,"failures"], sm[,"Walc"])
cor.test(sm1[,"failures"], sm1[,"Dalc"])
cor.test(sm1[,"failures"], sm1[,"Walc"])
cor.test(sm2[,"failures"], sm2[,"Dalc"])
cor.test(sm2[,"failures"], sm2[,"Walc"])
```

###goout
```{r}
cor.test(sm[,"goout"], sm[,"Dalc"])
cor.test(sm[,"goout"], sm[,"Walc"])
cor.test(sm1[,"goout"], sm1[,"Dalc"])
cor.test(sm1[,"goout"], sm1[,"Walc"])
cor.test(sm2[,"goout"], sm2[,"Dalc"])
cor.test(sm2[,"goout"], sm2[,"Walc"])
```

```{r}
sm <- student.mat

for (i in 1:32) {                                  
  sm[,i] <- as.factor(sm[,i])
}
```

###remi text
```{r}
layout(matrix(1:14, ncol = 7))
cdplot(sm[,"guardian"] ~ sm[,"Dalc"], data = sm, xlab = "Alcohol consumption during working days ", ylab = "Guardian", border = 1)
cdplot(sm[,"guardian"] ~ sm[,"Walc"], data = sm, xlab = "Alcohol consumption during Week-end ", ylab = "Guardian", border = 1)
cdplot(sm[,"Medu"] ~ sm[,"Dalc"], data = sm, xlab = "Alcohol consumption during working days ", ylab = "Father's education", border = 1)
cdplot(sm[,"Medu"] ~ sm[,"Walc"], data = sm, xlab = "Alcohol consumption during Week-end ", ylab = "Father's education", border = 1)
cdplot(sm[,"Fedu"] ~ sm[,"Dalc"], data = sm, xlab = "Alcohol consumption during working days ", ylab = "Mother's education", border = 1)
cdplot(sm[,"Fedu"] ~ sm[,"Walc"], data = sm, xlab = "Alcohol consumption during working Week-end ", ylab = "Mother's education", border = 1)
cdplot(sm[,"Pstatus"] ~ sm[,"Dalc"], data = sm, xlab = "Alcohol consumption during working days ")
cdplot(sm[,"Pstatus"] ~ sm[,"Walc"], data = sm, xlab = "Alcohol consumption during Week-end ")
cdplot(sm[,"famsize"] ~ sm[,"Dalc"], data = sm, xlab = "Alcohol consumption during working days ")
cdplot(sm[,"famsize"] ~ sm[,"Walc"], data = sm, xlab = "Alcohol consumption during working Week-end ")
cdplot(sm[,"Mjob"] ~ sm[,"Dalc"], data = sm, xlab = "Alcohol consumption during working days")
cdplot(sm[,"Mjob"] ~ sm[,"Walc"], data = sm, xlab = "Alcohol consumption during working Week-end")
cdplot(sm[,"famrel"] ~ sm[,"Dalc"], data = sm, xlab = "Alcohol consumption during working days")
cdplot(sm[,"famrel"] ~ sm[,"Walc"], data = sm, xlab = "Alcohol consumption during working Week-end ")
```

###remi text
```{r}
layout(matrix(1:10, ncol = 5))
cdplot(sm[,"romantic"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"romantic"] ~ sm[,"Walc"], data = sm)
cdplot(sm[,"goout"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"goout"] ~ sm[,"Walc"], data = sm)
cdplot(sm[,"internet"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"internet"] ~ sm[,"Walc"], data = sm)
cdplot(sm[,"activities"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"activities"] ~ sm[,"Walc"], data = sm)
cdplot(sm[,"freetime"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"freetime"] ~ sm[,"Walc"], data = sm)
```

###remi text
```{r}
layout(matrix(1:8, ncol = 4))
cdplot(sm[,"failures"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"failures"] ~ sm[,"Walc"], data = sm)
cdplot(sm[,"schoolsup"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"schoolsup"] ~ sm[,"Walc"], data = sm)
cdplot(sm[,"studytime"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"studytime"] ~ sm[,"Walc"], data = sm)
cdplot(sm[,"G3"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"G3"] ~ sm[,"Walc"], data = sm)

##Some regression
```{r}
RegFails <- lm (Walc ~ failures, data = sm)
summary (lm (Walc ~ failures, data = sm))
libraryr(ggplot2)
f <- ggplot(data =sm, aes(x = sm[,"Walc"], y = sm[,"failures"]))
f + geom_smooth(method = lm)
libraryr(ggplot2)
f <- ggplot(data =sm, aes(x = sm[,"Dalc"], y = sm[,"failures"]))
f + geom_smooth(method = lm)

sm1 <- subset(sm, Walc < 3 & Dalc <3)
sm2 <- subset(sm, Walc > 2 & Dalc > 2)

layout(matrix(1:2, ncol = 2))
cdplot(sm[,"failures"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"failures"] ~ sm[,"Walc"], data = sm)
```

#T
###GoOUT
```{r}
sm[,"goout"] <- as.numeric(sm[,"goout"])
sm[,"Dalc"] <- as.numeric(sm[,"Dalc"])
sm[,"Walc"] <- as.numeric(sm[,"Walc"])
cor.test(sm[,"goout"], sm[,"Dalc"])
cor.test(sm[,"goout"], sm[,"Walc"]) #Stong correlation
RegGoout <- lm (goout ~ Walc, data = sm)
RegGoout
libraryr(ggplot2)
f <- ggplot(data =sm, aes(x = sm[,"goout"], y = sm[,"Walc"]))
f + geom_smooth(method = lm) #TBS
sm[,"goout"] <- as.factor(sm[,"goout"])
sm[,"Dalc"] <- as.factor(sm[,"Dalc"])
sm[,"Walc"] <- as.factor(sm[,"Walc"])
layout(matrix(1:2, ncol = 2))
cdplot(sm[,"goout"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"goout"] ~ sm[,"Walc"], data = sm)

###Internet
```{r}
sm[,"internet"] <- as.numeric(sm[,"internet"])
sm[,"Dalc"] <- as.numeric(sm[,"Dalc"])
sm[,"Walc"] <- as.numeric(sm[,"Walc"])
cor.test(sm[,"internet"], sm[,"Dalc"])
cor.test(sm[,"internet"], sm[,"Walc"])
```

###Activities
```{r}
sm[,"activities"] <- as.numeric(sm[,"activities"])
sm[,"Dalc"] <- as.numeric(sm[,"Dalc"])
sm[,"Walc"] <- as.numeric(sm[,"Walc"])
cor.test(sm[,"activities"], sm[,"Dalc"])
cor.test(sm[,"activities"], sm[,"Walc"])

###Freetime
```{r}
sm[,"freetime"] <- as.numeric(sm[,"freetime"])
sm[,"Dalc"] <- as.numeric(sm[,"Dalc"])
sm[,"Walc"] <- as.numeric(sm[,"Walc"])
cor.test(sm[,"freetime"], sm[,"Dalc"])
cor.test(sm[,"freetime"], sm[,"Walc"])
```

```{r}
RegAbs <- lm (absences ~ Walc, data = sm)
RegAbs
library(ggplot2)
f <- ggplot(data =sm, aes(x = Walc, y = absences))
f + geom_smooth(method = lm) #TBS
sm[,"absences"] <- as.factor(sm[,"absences"])
sm[,"Dalc"] <- as.factor(sm[,"Dalc"])
sm[,"Walc"] <- as.factor(sm[,"Walc"])
layout(matrix(1:2, ncol = 2))
cdplot(sm[,"absences"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"absences"] ~ sm[,"Walc"], data = sm)
```

##Mean tests
```{r}
mean(sm[sm[, "Fedu"] == "5", "Dalc"]) - mean(sm[sm[, "Medu"] == "5", "Dalc"])
mean(sm[sm[, "guardian"] == "mother", "Dalc"]) - mean(sm[sm[, "guardian"] == "father", "Dalc"])
